---
title: "Plot Rivers Inlet Sockeye Salmon State-Space Spawner-Recruitment Model Results"
author: "B. Connors"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    collapsed: no
    fig_caption: yes
    highlight: espresso
    number_sections: yes
    smooth_scroll: yes
    theme: sandstone
    toc: yes
    toc_float: yes
---
# Purpose

The purpose of this document is to summarize spawner-recruitment model output and diagnostics in a unified format for ease of review. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE, cache=FALSE, include=TRUE)
options(scipen=1, digits=4)
# options(width=50, width.cutoff=50, digits = 6) 
require(kintr)
require(tidyverse)
require(PBSmodelling)
require(MASS)
require(RColorBrewer)
require(stringr)
require(lubridate)
require(pander)
require(kableExtra)
require(ggthemes)
require(xlsx)
require(plotly)
require(rstan)
require(tidybayes)
require(mcmcplots)
require(mltools)
require(viridis)

# Control Section
do.traceplots <- TRUE

# Read in Model Data and Output Files
stan.fit <- readRDS(file="./output/stan.fit.v2.rds")
stan.data <- readRDS(file="./output/stan.data.v2.rds")

fy <- 1948 # first year of spawner observations
years <- seq(fy, (fy+stan.data$nyrs-1))
brood_yrs <- seq(fy,fy+stan.data$nyrs-4)

stan.summary <- data.frame(summary(stan.fit)$summary)

pars <- rstan::extract(stan.fit)
```

# Basic Convergence Diagnostics
## Effective Sample Size 
Ideally effective sample sizes are > 400.

```{r}
hist(stan.summary$n_eff, 
     col="red", 
     breaks=50, 
     main="",
     yaxt="n",
     xlab="Effective sample size")
axis(2,las=2)
box(col="grey")
```

## R-hat Convergence Diagnostic
Compares the between- and within-chain estimates for model parameters and other univariate quantities of interest. If chains have not mixed well (i.e., the between- and within-chain estimates don't agree), R-hat is larger than 1. We recommend running at least four chains by default and only using the sample if R-hat is less than 1.05. Stan reports R-hat which is the maximum of rank normalized split-R-hat and rank normalized folded-split-R-hat, which works for thick tailed distributions and is sensitive also to differences in scale.

```{r}
hist(stan.summary$Rhat, 
     col="royalblue", 
     breaks=50, 
     main="",
     yaxt="n",
     xlab="R-hat")
axis(2,las=2)
box(col="grey")
```

# Key posterior estimates
Biological reference points are in hundreds of thousands and log-normal bias corrected
```{r}
kable(round(summary(stan.fit, pars= c("lnalpha", "lnalpha_c","beta", "sigma_R", "sigma_R0", "phi", "S_max", "S_eq_c", "S_msy_c", "S_msy_c","U_msy_c"))$summary,digits=2)[,c(1,4,6,8:10)],
      align = "crr") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

# Model Predictions and Fits
## Escapement 

Observations are blue points and model predictions are median (black line) and 50th and 97.5th intervals
```{r}
# Extract Data
obs <- stan.data$S_obs

# Model predictions
pred <-pars$S
pred.quant <- apply(pred, 2, quantile, probs=c(0.025,0.25,0.5,0.75,0.975))

spawners <- as.data.frame(cbind(years, obs, t(pred.quant)))
colnames(spawners) <- c("year","obs","lwr","midlwr","mid","midupper","upper")


# Plot
ggplot(spawners, aes(x=year, y = mid), show.legend = F) +
  geom_line(show.legend = F) + 
  geom_ribbon(aes(ymin = lwr, ymax = upper), show.legend = F, alpha=0.2) +
  geom_ribbon(aes(ymin = midlwr, ymax = midupper), show.legend = F, alpha=0.2) +
  geom_point(aes(x=years, y = obs), size=2,show.legend = F, col="royalblue") +
  xlab("Year") +
  ylab("Escapement (10^5) ") +
  theme(legend.position = "none") +
  theme_bw()
ggsave("./figures/escape-fit.jpg",height = 4.5, width = 6)

```

## Harvest 
Observations are blue points and model predictions are median (black line) and 50th and 97.5th intervals
```{r}
# Extract Data
obs <- stan.data$H_obs

# Model predictions
pred <-pars$C
pred.quant <- apply(pred, 2, quantile, probs=c(0.025,0.25,0.5,0.75,0.975))

spawners <- as.data.frame(cbind(years, obs, t(pred.quant)))
colnames(spawners) <- c("year","obs","lwr","midlwr","mid","midupper","upper")


# Plot
ggplot(spawners, aes(x=year, y = mid), show.legend = F) +
  geom_line(show.legend = F) + 
  geom_ribbon(aes(ymin = lwr, ymax = upper), show.legend = F, alpha=0.2) +
  geom_ribbon(aes(ymin = midlwr, ymax = midupper), show.legend = F, alpha=0.2) +
  geom_point(aes(x=years, y = obs), size=2,show.legend = F, col="royalblue") +
  xlab("Year") +
  ylab("Harvest (10^5) ") +
  theme(legend.position = "none") +
  theme_bw()
ggsave("./figures/harvest-fit.jpg",height = 4.5, width = 6)

```

## Escapement age composition 
Observations are blue points and model predictions are median (black line) and 50th and 97.5th intervals.
```{r}
# extract Data
obs <- stan.data$S_comps/rowSums(stan.data$S_comps)

# model predictions
pred <-pars$S_pa
pred.quant <- apply(pred, c(2,3), quantile, probs=c(0.025,0.25,0.5,0.75,0.975))

# wrangle observations and predictions into a data frame for ggplot
age_comps <- matrix(NA,(length(years)*4), 8)
colnames(age_comps) <- c("years","age_class","obs","lwr","midlwr","mid","midupper","upper")
age_comps[,1] <- rep(years,4)
age_comps[,2] <- rep(4:7,each = length(years))
age_comps[,3] <- as.vector(obs[,1:4])
age_comps[,4] <- as.vector(pred.quant[1,1:length(years),])
age_comps[,5] <- as.vector(pred.quant[2,1:length(years),])
age_comps[,6] <- as.vector(pred.quant[3,1:length(years),])
age_comps[,7] <- as.vector(pred.quant[4,1:length(years),])
age_comps[,8] <- as.vector(pred.quant[5,1:length(years),])
age_comps <- as.data.frame(age_comps)
age_comps$age_classF <- as.factor(age_comps$age_class)

# plot
ggplot(age_comps, aes(x=years, y = mid), show.legend = F) +
  geom_line(show.legend = F) + 
  geom_ribbon(aes(ymin = lwr, ymax = upper), show.legend = F, alpha=0.2) +
  geom_ribbon(aes(ymin = midlwr, ymax = midupper), show.legend = F, alpha=0.2) +
  geom_point(aes(x=years, y = obs), size=2,show.legend = F, col="royalblue") +
  facet_wrap(~age_classF, scales = "free_y")+
  xlab("Year") +
  ylab("Proportion") +
  theme(legend.position = "none") +
  theme_bw()
ggsave("./figures/esc-age-comp-fit.jpg",height = 4.5, width = 6)

```
## Harvest age composition 
Observations are blue points and model predictions are median (black line) and 50th and 97.5th intervals.
```{r}
# extract Data
obs <- stan.data$H_comps/rowSums(stan.data$H_comps)

# model predictions
pred <-pars$C_pa
pred.quant <- apply(pred, c(2,3), quantile, probs=c(0.025,0.25,0.5,0.75,0.975))
pred.quant[,49:72,] <- 0
# wrangle observations and predictions into a data frame for ggplot
age_comps <- matrix(NA,(length(years)*4), 8)
colnames(age_comps) <- c("years","age_class","obs","lwr","midlwr","mid","midupper","upper")
age_comps[,1] <- rep(years,4)
age_comps[,2] <- rep(4:7,each = length(years))
age_comps[,3] <- as.vector(obs[,1:4])
age_comps[,4] <- as.vector(pred.quant[1,1:length(years),])
age_comps[,5] <- as.vector(pred.quant[2,1:length(years),])
age_comps[,6] <- as.vector(pred.quant[3,1:length(years),])
age_comps[,7] <- as.vector(pred.quant[4,1:length(years),])
age_comps[,8] <- as.vector(pred.quant[5,1:length(years),])
age_comps <- as.data.frame(age_comps)
age_comps$age_classF <- as.factor(age_comps$age_class)

# plot
ggplot(age_comps, aes(x=years, y = mid), show.legend = F) +
  geom_line(show.legend = F) + 
  geom_ribbon(aes(ymin = lwr, ymax = upper), show.legend = F, alpha=0.2) +
  geom_ribbon(aes(ymin = midlwr, ymax = midupper), show.legend = F, alpha=0.2) +
  geom_point(aes(x=years, y = obs), size=2,show.legend = F, col="royalblue") +
  facet_wrap(~age_classF, scales = "free_y")+
  xlab("Year") +
  ylab("Proportion") +
  theme(legend.position = "none") +
  theme_bw()
```


## Spawner-recruitment relationship
Predictions are median (black line and points) and 97.5th intervals. Includes recruitment for incomplete brood years at end of time series. 

```{r,warning = FALSE,}
spwn <-pars$S
spwn.quant <- apply(spwn, 2, quantile, probs=c(0.025,0.5,0.975))[,1:(length(years)-4)]

rec <-pars$R
rec.quant <- apply(rec, 2, quantile, probs=c(0.025,0.5,0.975))[,8:dim(pars$R)[2]]


brood_t <- as.data.frame(cbind(years[1:(length(years)-4)],t(spwn.quant), t(rec.quant)))
colnames(brood_t) <- c("BroodYear","S_lwr","S_med","S_upr","R_lwr","R_med","R_upr")

brood_t <- as.data.frame(brood_t)

# SR relationship
spw <- seq(0,max(brood_t[,4]),length.out=100)
SR_pred <- matrix(NA,100,10000)
posteriors <- as.matrix(stan.fit)

for(i in 1:10000){
  r <- sample(seq(1,1000),1,replace=T)
  a <- pars$lnalpha_c[r]
  b <- pars$beta[r]
  SR_pred[,i] <- (exp(a)*spw*exp(-b*spw))
}

SR_pred <- cbind(spw,t(apply(SR_pred,c(1),quantile,probs=c(0.05,0.5,0.95),na.rm=T)))
colnames(SR_pred) <- c("Spawn", "Rec_lwr","Rec_med","Rec_upr")
SR_pred <- as.data.frame(SR_pred)
# 
# ggplot() +
#   geom_ribbon(data = SR_pred, aes(x = Spawn, ymin = Rec_lwr, ymax = Rec_upr), 
#               fill = "grey80", alpha=0.5, linetype=2, colour="gray46") +
#   geom_line(data = SR_pred, aes(x = Spawn, y = Rec_med), color="black", size = 1) +
#   geom_errorbar(data = brood_t, aes(x= S_med, y = R_med, ymin = R_lwr, ymax = R_upr), 
#                 colour="grey", width=0.2, size=0.3) +
#   geom_errorbarh(data = brood_t, aes(x= S_med, y = R_med, xmin = S_lwr, xmax = S_upr), 
#                  height=0, colour = "grey", width=0.2, size = 0.3) + 
#   geom_point(data = brood_t, aes(x = S_med, y = R_med, color=BroodYear, width = 0.9), size = 3)+
#   coord_cartesian(xlim=c(0, max(brood_t[,4])), ylim=c(0,max(brood_t[,7]))) +
#   scale_colour_viridis_c()+
#   xlab("Spawners (10^5)") +
#   ylab("Recruits (10^5)") +
#   theme_bw() +
#   theme(strip.text.x = element_text(size=8),
#         axis.title = element_text(size=10),
#         axis.text = element_text(size=7))+
#   geom_abline(intercept = 0, slope = 1,col="dark grey")
# ggsave("./figures/spawn-recuit.jpg",height = 4.5, width = 6)



p<- ggplot() +
  geom_ribbon(data = SR_pred, aes(x = Spawn, ymin = Rec_lwr, ymax = Rec_upr), 
              fill = "grey80", alpha=0.5, linetype=2, colour="gray46") +
  geom_line(data = SR_pred, aes(x = Spawn, y = Rec_med), color="black", size = 1) +
  geom_errorbar(data = brood_t, aes(x= S_med, y = R_med, ymin = R_lwr, ymax = R_upr), 
                colour="grey", width=0, size=0.3) +
  geom_errorbarh(data = brood_t, aes(x= S_med, y = R_med, xmin = S_lwr, xmax = S_upr), 
                 height=0, colour = "grey", height = 0, size = 0.3) + 
  geom_point(data = brood_t, aes(x = S_med, y = R_med, color=BroodYear, width = 0.9), size = 3)+
  coord_cartesian(xlim=c(0, 32), ylim=c(0,max(brood_t[,7]))) +
  scale_colour_viridis_c()+
  xlab("Spawners (10^5)") +
  ylab("Recruits (10^5)") +
  theme_bw() +
  theme(strip.text.x = element_text(size=8),
        axis.title = element_text(size=10),
        axis.text = element_text(size=7),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())+
  geom_abline(intercept = 0, slope = 1,col="dark grey")

ggExtra::ggMarginal(p + 
                      theme(legend.position = c(.85,.8),
                            legend.text = element_text(size = 7),
                            legend.title = element_text(size = 7),
                            legend.key.size = unit(0.5, 'cm')), 
                    fill="light grey", 
                    type = "histogram", 
                    colour = "grey")
ggsave("./figures/spawn-recuit.jpg",height = 5, width = 5.5)


```


## Spawner-recruitment relationship at low spawner abundances
```{r}

ggplot() +
  geom_ribbon(data = SR_pred, aes(x = Spawn, ymin = Rec_lwr, ymax = Rec_upr), 
              fill = "grey80", alpha=0.5, linetype=2, colour="gray46") +
  geom_line(data = SR_pred, aes(x = Spawn, y = Rec_med), color="black", size = 1) +
  geom_errorbar(data = brood_t, aes(x= S_med, y = R_med, ymin = R_lwr, ymax = R_upr), 
                colour="grey", width=0, size=0.3) +
  geom_errorbarh(data = brood_t, aes(x= S_med, y = R_med, xmin = S_lwr, xmax = S_upr), 
                 colour = "grey", height=0, size = 0.3) + 
  geom_point(data = brood_t, aes(x = S_med, y = R_med, color=BroodYear, width = 0.9), size = 3)+
  coord_cartesian(xlim=c(0, 10), ylim=c(0,20)) +
  scale_colour_viridis_c()+
  xlab("Spawners (10^5)") +
  ylab("Recruits (10^5)") +
  theme_bw() +
  theme(strip.text.x = element_text(size=8),
        axis.title = element_text(size=10),
        axis.text = element_text(size=7),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())+
  geom_abline(intercept = 0, slope = 1,col="dark grey")
ggsave("./figures/zoom-spawn-recuit.jpg",height = 5, width = 6.5)

```

## Productivity over time 
Predictions are median (black line) and 50th and 97.5th intervals.
```{r}

pred <-pars$lnalpha_y
pred.quant <- apply(pred, 2, quantile, probs=c(0.025,0.25,0.5,0.75,0.975))[,8:dim(pars$lnalpha_y)[2]]

log_prod_t <- as.data.frame(cbind(years[1:(length(years)-4)], t(pred.quant)))
colnames(log_prod_t) <- c("year","lwr","midlwr","mid","midupper","upper")

# Plot
ggplot(log_prod_t, aes(x=year, y = mid), show.legend = F) +
  geom_line(show.legend = F) + 
  geom_ribbon(aes(ymin = lwr, ymax = upper), show.legend = F, alpha=0.2) +
  geom_ribbon(aes(ymin = midlwr, ymax = midupper), show.legend = F, alpha=0.2) +
  xlab("Brood year") +
  ylab("Productivity (log) ") +
  theme(legend.position = "none") +
  theme_bw() + 
  geom_abline(intercept = 0, slope = 0 ,col="dark grey", lty=2)

ggsave("./figures/time-vary-prod.jpg",height = 4.5, width = 6)

```

# Biological benchmarks and stock-status.

```{r}
# Sgen1 function

"NLL_Sgen1" <- function(estSgen,aa,b,Smsy){	#minimize this function to get stock size that can recover to Smsy in one generation
	pSmsy=estSgen*exp(aa-b*estSgen)
	NLL=(log(pSmsy)-log(Smsy))^2
	return(NLL)
}

# Benchmarks
bench <- matrix(NA,10000,7,
                dimnames = list(seq(1:10000),c("lower_bm","upper_bm","recent_abund","recent_er","red","amber","green")))
posteriors <- as.matrix(stan.fit)

for(i in 1:10000){
  r <- sample(seq(1,1000),1,replace=T)
  a <- pars$lnalpha_c[r]
  b <- pars$beta[r]
  smsy <- pars$S_msy[r]
  fit=optim(par=(smsy*0.5),
            fn=NLL_Sgen1,
            aa=a ,
            b=b,
            Smsy=smsy,
            method="L-BFGS-B",
            hessian=F,
            lower=1,
            upper=smsy,
            control=list(fnscale=1))
  bench[i,1] <- fit$par
  bench[i,2] <- 0.8*smsy
  bench[i,3] <- mean(pars$S[r,(length(years)-6):length(years)])
  bench[i,4] <- mean(pars$U[r,(length(years)-6):length(years)])
  bench[i,5:7] <- 0
    if(bench[i,3]>= bench[i,2]){bench[i,7] <- 1}
  else{
    if(bench[i,3]< bench[i,1]){bench[i,5] <- 1}
    else{bench[i,6] <- 1}
  }

}

bench.quant <- apply(bench, 2, quantile, probs=c(0.025,0.25,0.5,0.75,0.975), na.rm=T)
sum(bench[,5],na.rm=T)/10000
sum(bench[,6],na.rm=T)/10000
sum(bench[,7],na.rm=T)/10000

bench_long <- pivot_longer(data = as.data.frame(bench[,1:3]),
             cols = c("lower_bm","upper_bm","recent_abund"),
             names_to = "metric",
             values_to = "value")

sub_bench_long <- subset(bench_long, value < 20)

ggplot(sub_bench_long, aes(x=value, fill=metric)) + 
  geom_histogram(alpha=0.6, position="identity", bins = 150)+
  xlab("Spawners (^5)") +
  ylab("Count") + 
  scale_fill_manual(values=c("red", "black", "dark green")) +
  coord_cartesian(xlim=c(0, 7), ylim=c(0,2500)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  labs(fill = "Metric" )+
  geom_vline(xintercept=c(bench.quant[3,1],bench.quant[3,2],bench.quant[3,3]),
             color=c("red","dark green","black"),
             alpha=0.7,
             lwd=1)

ggsave("./figures/status.jpg",height = 4, width = 6.75)

```


# Model Parameters
Summary of all model parameters.

```{r}
kable(data.frame(stan.summary), align="crr") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```


